---
title: "Data Madness"
output: html_notebook
---


Context:
Typically e-commerce datasets are proprietary and consequently hard to find among publicly available data. However, The UCI Machine Learning Repository has made this dataset containing actual transactions from 2010 and 2011. The dataset is maintained on their site, where it can be found by the title "Online Retail".

Content:
"This is a transnational data set which contains all the transactions occurring between 01/12/2010 and 09/12/2011 for a UK-based and registered non-store online retail.The company mainly sells unique all-occasion gifts. Many customers of the company are wholesalers."

Acknowledgements:
Per the UCI Machine Learning Repository, this data was made available by Dr Daqing Chen, Director: Public Analytics group. chend '@' lsbu.ac.uk, School of Engineering, London South Bank University, London SE1 0AA, UK.

Inspiration:
Analyses for this dataset could include time series, clustering, classification and more.

```{r include=FALSE}
#Loading Libraries
library(data.table)
library(lubridate)
library(ggplot2) # Data visualization
library(dplyr)
library(DataExplorer)
library(forecast)
library(TSA)
```
```{r}
#Loading E-commerce data
df_data <- fread('data.csv')
head(df_data)
```
```{r}
plot_missing(df_data)
```
```{r}
df_data <- na.omit(df_data)
dim(df_data)
```
```{r}
df_data$date <- sapply(df_data$InvoiceDate, FUN = function(x) { strsplit(x, split="[ ]")[[1]][1]})
df_data$time <- sapply(df_data$InvoiceDate, FUN = function(x) { strsplit(x, split="[ ]")[[1]][2]})

df_data$date <- as.Date(df_data$date, "%m/%d/%Y")
df_data$dayOfWeek <- wday(df_data$date, label=TRUE)

df_data <- df_data %>% mutate(totalPrice = UnitPrice * Quantity)

head(df_data)
```
```{r}

revenueByDate <- df_data %>% 
    group_by(date) %>%
    summarise(revenue = sum(totalPrice))

ts_data <- ts(revenueByDate$revenue, start = c(2010,12), frequency = 305)
revenueByDate$clean_ts <- tsclean(ts_data)
ggplot() + 
  geom_line(data=revenueByDate, aes(x = date, y = clean_ts, color = "Volume")) + labs(x='Date', y='Revenue (£)', title='TS Revenue by Date')

df_data %>%
  group_by(date) %>%
  summarise(revenue = sum(totalPrice)) %>%
  ggplot(aes(x = date, y = revenue)) + geom_line() + geom_smooth(method="auto", se=TRUE) + labs(x='Date', y='Revenue (£)', title='Revenue by Date')

df_data %>% 
  group_by(dayOfWeek) %>%
  summarise(revenue = sum(totalPrice)) %>%
  ggplot(aes(x=dayOfWeek, y=revenue)) + geom_col() +
  labs(x='Day of Week', y='Revenue (£))',title='Revenue by day of the week')
```
```{r}
weekday_df <- df_data %>%
  group_by(date, dayOfWeek) %>%
  summarise(revenue = sum(totalPrice), transactions=n_distinct(InvoiceNo)) %>%
  mutate(aveRevenue = (round((revenue / transactions),2))) %>%
  ungroup()


head(weekday_df)

```
```{r}
ggplot(weekday_df, aes(x=dayOfWeek, y=revenue)) + geom_boxplot() +labs(x='Day of the week', y='Revenue (£)', title='Revenue by day of the week')
ggplot(weekday_df, aes(x=dayOfWeek, y=transactions)) + geom_boxplot() +labs(x='Day of the week', y='Transaction', title='Number of transactions by day of the week')
ggplot(weekday_df, aes(x=dayOfWeek, y=aveRevenue)) + geom_boxplot() +labs(x='Day of the week', y='Avg. Revenue', title='Avg revenue by day of the week')

```
#Country summary
```{r}
country_df <- df_data %>%
  group_by(Country) %>%
  summarise(revenue=sum(totalPrice), transactions=n_distinct(InvoiceNo)) %>%
  mutate(avgOrderVal = round((revenue / transactions),2)) %>%
  arrange(desc(revenue))

head(country_df)
```
```{r}
topFiveCountries <- df_data %>%
filter(Country == 'Netherlands' | Country == 'EIRE' | Country == 'Germany' | Country == 'France' | Country == 'Australia')

topFiveCountries_df <- topFiveCountries %>%
  group_by(Country, date) %>%
  summarise(revenue = sum(totalPrice), transactions= n_distinct(InvoiceNo), customers = n_distinct(CustomerID)) %>%
  mutate(avgOrdVal = (round((revenue/transactions), 2))) %>%
  ungroup() %>%
  arrange(desc(revenue))

head(topFiveCountries_df)
```
```{r}
ggplot(topFiveCountries_df, aes(x=Country, y=revenue)) + geom_col() + labs(x = 'Country', y='Revenue (£)', title='Revenue by Country')
ggplot(topFiveCountries_df, aes(x=date, y=revenue, colour=Country))+geom_smooth(method='auto', se=FALSE) + labs(x='Date',y='Revenue (£)') 
ggplot(topFiveCountries_df, aes(x=Country, y=avgOrdVal)) + geom_boxplot() + labs(x='Country', y='Average Order Value (£)') + scale_y_log10()
ggplot(topFiveCountries_df, aes(x=Country, y=transactions)) + geom_boxplot() + labs(x='Country', y="Transactions", title = "Number of transactions by Country")
```
```{r}
sales_by_stock <- df_data %>% 
  group_by(StockCode) %>%
  summarize(sales=sum(Quantity))


sales_by_stock[order(-sales_by_stock$sales),]
print ("The best selling item has stock code 84077. Description: ")
df_data[df_data$StockCode==84077,]$Description[1]

print ("The worst selling item has stock code 21700. Description: ")
df_data[df_data$StockCode==21700,]$Description[1]

```
```{r}
revenue_by_stock <- df_data %>% 
  group_by(StockCode) %>%
  summarize(revenue=sum(Quantity*UnitPrice))

revenue_by_stock[order(-revenue_by_stock$revenue),]
print ("The item which caused the highest revenue has stock code: 22423. Description:")
df_data[df_data$StockCode=="22423",]$Description[1]
print ("The item which caused the lowest revenue has stock code: 23359 Description:")
df_data[df_data$StockCode=="23359",]$Description[1]

```
```{r}
customer_spending <- df_data %>%
  group_by(CustomerID) %>%
  summarize(purchases = sum(Quantity*UnitPrice))

customer_spending[order(-customer_spending$purchases),]
print ("The customer spending the most has ID: 14646 and he purchased for £279489.02")
print ("The customer spending the least has ID: 15594 and he purchased for £1767.530")
```
```{r}
adf.test(revenueByDate$clean_ts, alternative = "stationary")
fit <- auto.arima(revenueByDate$clean_ts, seasonal = TRUE)
fit
tsdisplay(residuals(fit), lag.max=45, main='ARIMA(5,1,0) ')
fcast <- forecast(fit, h=30)
plot(fcast)
```
```{r}
x.spec <- spec.pgram(revenueByDate$clean_ts,log="no", taper=0)
spx <- x.spec$freq*1 #to get the correct x axis (note that this is only approximation - see example below for accurate)
plot(x.spec$spec~spx,xlab="period",ylab="spectral density",type="h")
title("E-commerce periodogram")
```

